---
title: "Final Report"
author: "Juliana Bushnell"
date: "10/23/2022"
output: html_document
---
Every year, the NFL puts on a ‘data bowl’ for armature sports analysts. In 2020, their challenge was using player tracking data to predict how many yards will an NFL player gain after receiving a handoff. There are 49 variables and 682154 observations. 

The purpose of this file is to analyze a few of the variables in order to understand the dataset more. As the semester continues, I will continue to add more to it. Initially, I am simply focused on understanding the distribution of rushing plays by team and understanding the player build. 


```{r, package load, echo=FALSE}
library(knitr)
library(readr)
library(kableExtra)
library(ggplot2)
library(dplyr, warn.conflicts = FALSE)
```


# Importing Data & Verify
```{r, message=FALSE, warning=FALSE, import, echo=FALSE}
here::i_am("final_report.Rmd")
train <- read_csv(file = here::here("data/train.csv"))
player<- read_csv(file=here::here("data/playerdata.csv"))
```

# Variable List
```{r, explain, echo=FALSE}
ls(train)
```

## Var Details 

### Game Information
GameId - a unique game identifier 
Season - year of the season 
HomeTeamAbbr - home team abbreviation  
VisitorTeamAbbr - visitor team abbreviation  
Week - week into the season  
Stadium - stadium where the game is being played  
Location - city where the game is being played  
StadiumType - description of the stadium environment  
Turf - description of the field surface  
GameWeather - description of the game weather  
Temperature - temperature (deg F)  
Humidity - humidity  
WindSpeed - wind speed in miles/hour  
WindDirection - wind direction  

### Player Information
NflId - a unique identifier of the player 
DisplayName - player's name  
JerseyNumber - jersey number 
PlayerHeight - player height (ft-in)  
PlayerWeight - player weight (lbs)  
PlayerBirthDate - birth date (mm/dd/yyyy)  
PlayerCollegeName - where the player attended college  
Position - the player's position (the specific role on the field that they typically play)  

### Play Information
PlayId - a unique play identifier  
X - player position along the long axis of the field. See figure below.  
Y - player position along the short axis of the field. See figure below. 
S - speed in yards/second  
A - acceleration in yards/second^2  
Dis - distance traveled from prior time point, in yards  
Orientation - orientation of player (deg)  
Dir - angle of player motion (deg)  
YardLine - the yard line of the line of scrimmage  
Quarter - game quarter (1-5, 5 == overtime)  
GameClock - time on the game clock  
PossessionTeam - team with possession  
Down - the down (1-4)  
Distance - yards needed for a first down  
FieldPosition - which side of the field the play is happening on  
HomeScoreBeforePlay - home team score before play started  
VisitorScoreBeforePlay - visitor team score before play started 
NflIdRusher - the NflId of the rushing player  
OffenseFormation - offense formation  
OffensePersonnel - offensive team positional grouping  
DefendersInTheBox - number of defenders lined up near the line of scrimmage, spanning the width of the offensive line  
DefensePersonnel - defensive team positional grouping  
PlayDirection - direction the play is headed  
TimeHandoff - UTC time of the handoff  
TimeSnap - UTC time of the snap  
Yards - the yardage gained on the play (you are predicting this)  
Team - home or away  

## Details 
The first year in this data set is `r min(train$Season)` and goes until `r max(train$Season)`
This tables shows the amount of recorded rushing plays per team per season by using the count of teams in possession during a play 
``` {r, rushing play table, echo=FALSE}
knitr::include_graphics(
  here::here("figures/output1.png")
)

```

## Plot 

This plot shows the distribution of player weights within positions. The range of player weight is `r min(train$PlayerWeight)` to `r max(train$PlayerWeight)`. This graph doesn't have to do with the root of the dataset which is trying to understand rushing plays. Rushing players are typically carried out by running backs, quarterbacks, and wide receivers. This graph illustrates in general the weight of players between the 2017-2019 season. 
``` {r, Player Weight among positions}
knitr::include_graphics(
  here::here("figures/height_weight.png")
)

```

##Table 3: Distribution of Colleges among positions
This table shows the amount of rushing plays under different weather conditions by week.  
``` {r, positions by colleges}
positioncollege <- table(player$position, player$playercollegename)
library(knitr)
library(webshot)
positioncollege  %>%
  kbl() %>%
  kable_paper("hover", full_width = F) 

```

